---
title: "Workshop 5"
author: "Kaysha Kenney"
date: "2023-05-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load into R library 
```{r}
library(tidyverse)
library(sf) # simple features
library (terra) # for raster
library(tmap) # Thematic maps are geographical maps in which spatial data distributions are visualized

```

# Upload copepod data set. Took the folder from my desktop and put it in my GitHub folder and into R and into my end to end analysis folder so the data set in now in my R, which I can then upload. 
```{r}
library(readr)
dat <- read_csv("data-for-course/copepods_raw.csv")
dat
  
```


# 6.7.1 Check coordinates ### 
# First we need to make a ggplot to plot the coordinates for the samples 

```{r}
library(ggplot2)
ggplot(dat) + 
  aes(x = longitude, y = latitude, color = richness_raw) +
  geom_point()
```

# Look at the richness data (main variable). Visualize richness in non-spatial way with latitue on the x axis and richness on the y axis.  
```{r}
ggplot(dat, aes(x = latitude, y = richness_raw)) + 
  stat_smooth() + 
  geom_point()
```

# 6.8 Getting going with maps ### 
# repeat the same richness but with packages for GIS and mapping. Turn data into 'simple features collection' 
```{r}
sdat <- st_as_sf(dat, coords = c("longitude", "latitude"), 
                 crs = 4326)
```

# 6.9 Coordinate Reference Systems ### 

```{r}
crs4326 <- st_crs(4326)
crs4326 # look at the whole CRS
crs4326$Name # pull out just the name of the crs
```
```{r}
crs4326$wkt # crs in well-known text format
```
# 6.10 Feature Collection ### 

```{r}
sdat
```
# 6.11 Cartography ### 

# Mapping, only plotting the richness column. 
```{r}
plot(sdat["richness_raw"])

```
```{r}
plot(sdat)
```

# 6.12 Thematic maps for communication ### 

# Going to build and add on layers 
```{r}
#using tmap
tm1 <- tm_shape(sdat) + 
  tm_dots(col = "richness_raw")
tm1
```

# Save the map to the working directory. Make sure to change the output path if I need to save it to a different folder 
```{r}
tmap_save(tm1, filename = "Richness-map.png", 
          width = 600, height = 600)
```

# 6.13 Mapping spatial polygons as layers ### 
# Read the shapefiles directly into R with the st_read comman, similar to read_csv but for spatial files 
```{r}
aus <- st_read("data-for-course/spatial-data/Aussie/Aussie.shp")
```
```{r}
shelf <- st_read("data-for-course/spatial-data/aus_shelf/aus_shelf.shp")
regions <- st_read("data-for-course/spatial-data/regions/regions.shp")
```
```{r}
aus
```

# 6.13.12 Mapping your polygons ### 

# Making a map of polygons 
```{r}
tm_shape(shelf) + 
  tm_polygons()
```
```{r}
tm_shape(shelf, bbox = sdat) + 
  tm_polygons() +
  tm_shape(aus) + 
  tm_polygons() + 
  tm_shape(sdat) + 
  tm_dots()
```

# 6.14 Exploring t_map ### 

# Changing the style of the map, similar to ggplot themes function, allows us to customize it 
```{r}
vignette('tmap-getstarted')
```

# Customizing the Australia map 
```{r}
tm2 <- tm_shape(regions) +
  tm_polygons(col = "region", palette = "magma") +
  tm_shape(shelf, bbox = sdat) + 
  tm_polygons() +
  tm_shape(aus) + 
  tm_polygons(col = "adm") + 
  tm_shape(sdat) + 
  tm_dots(col = "richness_raw")+ 
  tmap_style("natural")
```

```{r}
tm_shape(regions) +
  tm_polygons(col = "region", palette = "magma") +
  tm_shape(shelf, bbox = sdat) + 
  tm_polygons() +
  tm_shape(aus) + 
  tm_polygons(col = "adm") + 
  tm_shape(sdat) + 
  tm_dots(col = "richness_raw")+ 
  tmap_style("natural") 

```
# save the new customized map 
```{r}
tmap_save(tm2, filename = "Richness-map.png", 
          width = 600, height = 600)
```

